Title

### 2024-05-28 paper/main.tex
Updates

\begin{abstract}
	The initial step in a causal analysis involves determining the causal
	structure between the variables from data typically in the form of a
	Directed Acyclic Graph (DAG) or an Structural Equation Model (SEM).
	This procedure, known as \emph{causal discovery}, has been extensively
	studied for both DAGs and SEMs. While numerous automated algorithms
	exist for estimating DAGs from datasets, their application in applied
	fields remain limited, possibly due to a lack of confidence in these
	algorithms stemming from them making obvious mistakes, and the
	difficulties in choosing the right algorithm for a given dataset. As a
	result, in applied disciplines, DAGs are predominantly constructed
	manually based on domain expertise making it vital to test the fit of
	these models to the data and potentially modifying them to improve the
	fit. Conversely, SEMs are typically constructed manually where
	researchers come up with an initial model based on domain knowledge and
	then iteratively assess its fit and modify the model according to that.
	This process is commonly known as \emph{specification search} where
	expert knowledge combined with data-driven methods like
	\emph{modification indices} is used to guide the modification process.
	In the case of DAGs, although statistical tests are available that can
	assess the fit of the model, there are no methods that can guide our
	modification process to improve the fit.

	Inspired by the specification search in SEMs, in this paper, we
	introduce a data-driven modification process aimed at assisting
	researchers in manually constructing or modifying DAGs generated by an
	algorithm. The first step of this process involves identifying pairs of
	variables whose observed association is not fully explained by the
	model. The user can then make modifications to the model by adding
	edges between such variables while using their domain expertise to
	decide the direction of the edge. For the measure of association, in
	the case of continuous and discrete variables we can use effect sizes,
	however no such measure of association is present for mixed data. We
	propose a novel (conditional) measure of association for mixed data
	based on canonical correlations. We also present a graphical web-tool
	that can assist researchers in iteratively modifying and constructing
	their DAGs. We theoretically show that in the presence of an oracle,
	this iterative modification process is consistent in recovering the
	true DAG. Empirically, we show that this process performs comparably to
	PC and Hill-Climb Search algorithms if the user is able to correctly
	identify the direction of the edge in one out of three cases.
\end{abstract}

>>>>

\begin{abstract}
	The initial step in a causal analysis involves determining the causal
	structure between the variables from data typically in the form of a
	Directed Acyclic Graph (DAG) or an Structural Equation Model (SEM).
	This procedure, known as \emph{causal discovery}, has been extensively
	studied for both DAGs and SEMs. While numerous automated algorithms
	exist for estimating DAGs from datasets, their application in applied
	fields remain limited, possibly due to a lack of confidence in these
	algorithms stemming from them making obvious mistakes, and the
	difficulties in choosing the right algorithm for a given dataset.
	Consequently, in applied disciplines, these DAGs are predominantly
	constructed manually based on domain expertise. Due to being
	constructed by hand, it is vital that the fit of these models are
	tested against data and the models are potentially modified to acheive
	a better fit. On the other hand, in the field of SEMs models are
	typically constructed manually through an iterative process of
	assessing fit and modifying the model according to that. This process
	is commonly known as \emph{Specification Search} where methods like
	\emph{modification indices} combined with expert knowledge can be used
	to guide this modification process. In the case of DAGs, although there
	are tests to assess the global fit of a model by combining the tests
	using implied conditional independencies of the model, but there are no
	methods that can guide our modification process. In this paper, we
	present a simple modification process aimed at helping researchers in
	manually constructing or modifying DAGs outputted by an algorithm. This
	modification process is based on using a (conditional) measure of
	association between variables in the model to assess potential edges
	that best explains unexplained correlation between variables which the
	researcher specifies the direction of this edge. In the case of
	continuous or discrete variables, the effect size measures of local
	tests can be used as this measure of association, however for mixed
	data there is no such measure. We present a measure of association
	based on canonical correlations for mixed data. We also present a
	graphical web-tool that can assist researchers in iteratively modifying
	and constructing their DAGs. We theoretically show that in the presence
	of an oracle, this iterative modification process is consistent in
	recovering the true DAG. Emprically, we show that using the mixed data
	measure of association, the modification process performs comparably to
	PC and Hill-Climb Search algorithms if the user is able to correctly
	identify the direction of the edge in one out of three cases.
\end{abstract}
